---
title: "weighting_investigation_rapid_markets"
author: "Zack Arno"
date: "3/9/2021"
output:
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, message= F)

library(tidyverse)
library(srvyr)
library(glue)


df<- read_csv("outputs/20210310_clean_data.csv")
host_pop<- read_csv("inputs/20210224_host_pop_cleaned.csv")
ref_pop<- read_csv("inputs/ref_pops.csv")

dt_set_options<- function(x){
    DT::datatable(x,
    options = list(
      autoWidth=F,
      dom= 't',
     list(list(width = '20%', targets = list(1,2,3)))
     )
    )}

```

## Sampling Methodology 
Sampling for the february/march 2021 rapid market assessmet in Isingiro & Lamwo was done using stratified SRS for the refugee population and a variation of cluster sampling for the host community. As the cluster sampling method that was followed presents several limitations, I am attempting to use this markdown report to find the best method to weight the data and create the design objects for analysis under the current circumstances.

The **host community** "cluster" sampling method was as follows: the lowest admin level of population data available was sub-county. However, this was not used as the cluster. A cluster size of 6 was selected and the number of clusters drawn for each sub-county was based on the population of that sub-county with a target of 95-7 confidence/margion of error at the district level. The design effect to calculate total sample needed was taken from the 2017 UBOs National Household Assessment. The clusters themselves were randomly generate gps points. When enumerators arrived at the cluster pt location they were told to sample the 6 closest household

### Limitations
Some of the limitations to the cluster sampling: 

* no population data for the clusters (host community)
* a bias towards rural populations
* Since we have no population

### Steps to deal with/ minimize limitations

1. split the datasets by population (refugee, host)
2. weight refugee data  at district level and analyze as stratified SRS at the district level and then overall refugee level
3. weight host data at sub-county level and  analyze as clusters sample at the district and then overall host level (details below)
4. combine the two data sets as a cluster sample, weight at district level and aggregate both populations together at the district and overall level.

By creating several sets of weights and design objects I avoid having the limitations imposed by the host community sampling methodology impact the analysis of the refugee data (step 2). However, the limitations will still effect the analysis (step 4) when I analyze the refugee and host data together.


## 1.)  Refugee only weight & survey design construction

The weighting and methodology for the refugee only analysis (district & refugee overall)  can be seen clearly from the code and table below.

```{r refugee weight table}
# step 1 - calculate weights and make design ------------------------------

# need to make sure if these are hhs or individuals

# df %>% tabyl(status_loc)
# just some recoding
df<-df %>% 
  mutate(
    strata= case_when(
      status_loc %in% c("nakivale_refugee","oruchinga_refugee")~"isingiro_refugee",
      status_loc == "palabek_refugee"~ "lamwo_refugee",
      TRUE ~ status_loc
    )
  )

#filter to only refugee data
df_ref<- df %>% 
  filter(str_detect(strata, "_refugee$"))

# only hos
df_host<- df %>% 
  filter(str_detect(strata, "_host$"))

# some recoding of the refugee population 
ref_pop<- ref_pop %>% 
  mutate(strata= ifelse(strata=="Isingiro_ref",
                        "isingiro_refugee",
                        "lamwo_refugee"),
         pop_status= "refugee") %>% 
  rename(total_hh="total_pop")

# create weight table that can be merged with data to create a design object
refugee_weight_table<-df_ref %>% 
  group_by(strata) %>% 
  summarise(
    sample_strata=  n()
  )%>% 
  mutate(
    sample_global= sum(sample_strata)) %>% 
  left_join(ref_pop %>% 
              rename(pop_strata= "total_hh")) %>% 
  mutate(pop_global=sum(pop_strata) ,
  weights = (pop_strata/pop_global)/(sample_strata/sample_global)
  )

# refugee_weight_table
refugee_weight_table %>% dt_set_options()

```


### Refugee survey design object 

This survey design object can be used to analyze the refugee data per district and to aggregate the refugee data together

```{r, refugee design object}

df_ref<-df_ref %>% 
  left_join(refugee_weight_table, by= "strata")

refsvy<-survey::svydesign(ids = ~ 1,
                              strata =  ~strata,
                              weights= ~weights,
                              data = df_ref)

refsvy<- as_survey(refsvy)
# as_survey()
```


## 2.) Host community only weight & survey design construction

```{r, echo = F}
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
df_host<- df_host %>% 
  group_by(unique_pt_id) %>% 
  mutate(sub_county_name= mode(sub_county_name)) %>% 
  ungroup()

```

A table created below shows the host community population data at the sub-county level
```{r}


df_host<- df_host %>% 
    mutate(
      district_sub_county= paste0(district_name, "_",sub_county_name)
    )
  
  # some annoying recoding to make match data set
  
  host_pop_harmonized<- host_pop %>% 
    mutate(district_name= tolower(district),
           district_sub_county= glue::glue("{district}_{sub_county}"),
           district_sub_county = ifelse(district_sub_county=="isingiro_town_council",
                          "isingiro_isingiro_tc",district_sub_county),
           sub_county_name= ifelse(sub_county=="isingiro_town_council",
                              "isingiro_tc",sub_county)
           
    ) %>% 
    select(district_name,sub_county_name, district_sub_county, total_hh=hh)
 
  host_pop_harmonized %>% 
  dt_set_options()

```


```{r,big problem , eval=F, echo=F}

df_host %>%
  group_by(unique_pt_id) %>%
  mutate(num_sub_counties =length(unique(sub_county_name)),
            ) %>%
  filter(num_sub_counties!=1) %>%
  mutate(name= "sub_county_name",
         type= "change_response"
         current_value= sub_county_name,
         value= mode(sub_county_name),
         issue_id = "wrong_sub_county",
         adjust_log="",
         Reviewed="") %>%
  select(uuid, name, current_value, value, issue_id, adjust_log, Reviewed) %>%
  arrange(unique_pt_id)# %>%
  #write_csv("inputs/auto_cl/20210310_wrong_subcounties_autocl.csv")




```


```{r, eval=F, echo=F}
df_host %>% 
  group_by(unique_pt_id) %>% 
  mutate(
    n=n()
  ) %>% 
  filter(n!=6) %>% 
   mutate(name= "unique_pt_id",
         type= "change_response",
         current_value= unique_pt_id,
         value= "",
         issue_id = "cluster_size_not_6",
         adjust_log="",
         Reviewed="",
         clust_size= n) %>%
  select(uuid, name, current_value, value, issue_id, adjust_log, Reviewed, clust_size) %>%
  arrange(unique_pt_id) %>% 
  write_csv("inputs/auto_cl/20210310_cluster_not_6.csv")

# still a problem
df_host %>% 
  mutate(
    cluster_id= paste0(strata, "_", parse_number(unique_pt_id))
    ) %>% 
  group_by(cluster_id) %>% 
  summarise(
    sample_strata= n()
  ) %>% 
  filter(sample_strata!=6)

# doest this mean that each cluster is not the same sub_county?

```


To weight the clusters we need assume: 

*  Each cluster in the the same sub-county has equal populations and thus equal weights
  + I think we can therefore weight at the sub-county level with each cluster in the same sub-county having the same weight. I think this will be effectively the same thing as weighting at the district level.
  + Below I make two sets of weights to test this theory

```{r, echo=F, eval= F}
cluster_breakdown_table<-df_host %>% 
  mutate(
    cluster_id= paste0(strata, "_", parse_number(unique_pt_id))
    ) %>% 
  group_by(sub_county_name) %>% 
  summarise(number_clusters= length(unique(cluster_id)),
            number_interviews = n()
  )

cluster_breakdown_table %>%  dt_set_options()


```


**code below:** creating the weights by district
```{r, weighting by district}
census_pop_stats_by_district<- host_pop_harmonized %>% 
  group_by(district_name) %>% 
  summarise(
    pop_strata= sum(total_hh)
  ) %>% 
  ungroup() %>% 
  mutate(
    pop_global= sum(pop_strata)
  )

sample_pop_stats_by_district<-df_host %>% 
  group_by(district_name) %>% 
    summarise(
    sample_strata=  n()
  ) %>% 
  ungroup() %>% 
  mutate(
    sample_global = sum(sample_strata)
  ) 
host_district_weight_table<- census_pop_stats_by_district %>% 
  left_join(sample_pop_stats_by_district, by="district_name") %>% 
  mutate(
    district_weights= (pop_strata/pop_global)/(sample_strata/sample_global)
  )

```


**code below:** creating the weights by sub county
```{r, weighting by sub_county}
census_pop_stats_by_sub_county<- host_pop_harmonized %>% 
  group_by(sub_county_name) %>% 
  summarise(
    pop_strata= sum(total_hh)
  ) %>% 
  ungroup() %>% 
  mutate(
    pop_global= sum(pop_strata)
  )

sample_pop_stats_by_sub_county<-df_host %>% 
  group_by(sub_county_name) %>% 
    summarise(
    sample_strata=  n()
  ) %>% 
  ungroup() %>% 
  mutate(
    sample_global = sum(sample_strata)
  ) 
host_sub_county_weight_table<- census_pop_stats_by_sub_county %>% 
  left_join(sample_pop_stats_by_sub_county, by="sub_county_name") %>% 
  mutate(
    sub_county_weights= (pop_strata/pop_global)/(sample_strata/sample_global)
  )

```


**code below:** adding both weights to the data set
```{r adding weights}

df_host_with_weights<-df_host %>% 
  mutate(
    cluster_id= paste0(strata, "_", parse_number(unique_pt_id))
  ) %>% 
  left_join(
    host_district_weight_table %>% select(district_name, district_weights), by ="district_name"
  ) %>% 
  left_join(
    host_sub_county_weight_table %>%  select(sub_county_name, sub_county_weights), by= "sub_county_name"
  )


# host_weight_table %>% dt_set_options()
```

Below is a table showing the data with a few relevant columns including different weights. This is shown to verify that weights were set up as intended. Looks good

```{r eval=F, echo=F}

# problem if there is only one cluster in  strata... .thats the problem 
# if we make the sub-county the strata this happens... if we make the the strata the district
# then this will not be an issue.....
# 
# problem_cluster_id<- df_host_with_weights %>% 
#   group_by(cluster_id) %>% 
#   summarise(n=n()) %>% 
#   filter(n<3) %>% pull(cluster_id)
# dfhost2<- df_host_with_weights %>%
#   filter(cluster_id!=problem_cluster_id & !str_detect(!!sym(host_strata_level),"kakamba"))
```

```{r, showing weights and data}
df_host_with_weights %>% 
  select(cluster_id, district_name,sub_county_name,  sub_county_weights, district_weights) %>% 
  head(50) %>% 
  dt_set_options()
```

### Host survey design object

**code below:** create two host community only survey design objects. I have made one survey object using the district level weights and one using the sub_county level weights.

```{r host svy design object}

hostsvy_dist <- survey::svydesign(data = df_host_with_weights,
                             ids = ~cluster_id, 
                             strata = ~district_name,
                             weights = ~district_weights
                             # nest = T
                               )
hostsvy_dist<- as_survey(hostsvy_dist)

hostsvy_subc <- survey::svydesign(data = df_host_with_weights,
                             ids = ~cluster_id, 
                             strata = ~district_name,
                             weights = ~sub_county_weights
                             # nest = T
                               )
hostsvy_subc<- as_survey(hostsvy_subc)
```


**code below:** Calculate % mean of random categorical variable using  district weights

```{r}
hostsvy_dist %>% 
  group_by(crop_growing) %>% 
  summarise(
    pct= survey_mean( vartype="ci")
    ) %>% dt_set_options()
```

**code below:** Calculate % mean of random categorical variable using  sub county weights. 

```{r}
hostsvy_subc %>% 
  group_by(crop_growing) %>% 
  summarise(
    pct= survey_mean( vartype="ci")
    ) %>% dt_set_options()

```
Nice, they are the same

## 3.) Host Community  & Refugee combined weight & survey design construction

For aggregating host community and refugee together at the district and overall levels I need to define one cluster survey design. Since I have just shown that sub-county and district level weights produce the same result when aggregated to the district level I will use district weights


**code below:**  construct the weights and add them to the data set
```{r,eval=T}

weight_tables<- list(refugee_weight_table, host_district_weight_table %>% 
                       rename(strata= "district_name"))


weight_table<-map_dfr(weight_tables, function(x)x %>% 
          select(pop_group=strata, sample_strata, pop_strata)) %>% 
  mutate(pop_group= ifelse(!str_detect(pop_group,"_refugee$"),glue::glue("{pop_group}_host"),pop_group))

weight_table<-  weight_table %>% 
  mutate(
    pop_global= sum(pop_strata),
    sample_global = sum(sample_strata),
    weights = (pop_strata/pop_global)/(sample_strata/sample_global)
  )

#
df<- df %>% 
  mutate(
     cluster_id= case_when(
       status=="host_community"~
       paste0(strata, "_", parse_number(unique_pt_id)),
       TRUE ~strata),
    pop_group= case_when(
      status=="host_community"~glue("{district_name}_host"),
      TRUE~ strata
                          )
  )


df<- df %>% 
  left_join(
    weight_table, by =c("pop_group")
  )

```



**code below:** make the surey object and calculate the same mean % as above, but this time for the total population
```{r}
df<-df %>% 
  mutate(
    overall= "overall"
  )
host_ref_svy <- survey::svydesign(data = df,
                             ids = ~cluster_id, 
                             strata = ~ district_name, 
                             weights = ~weights,
                             nest = T
                             )
host_ref_svy<- as_survey(host_ref_svy)


host_ref_svy %>% 
  group_by(crop_growing) %>% 
  summarise(
    pct= survey_mean( vartype="ci")
    ) %>% dt_set_options()

```




```{r,eval=F, echo= F}
weights = (55/pop_global)/(55/sample_global)

sg/popg
55/pg * sg/55
```


```{r,eval=F, echo=F}
host_pop$strat[host_pop$strat %in% df_host$sub_county_name]
refugee_weight_table<-df_host %>% 
  group_by(strata) %>% 
  summarise(
    sample_strata=  n()
  )%>% 
  mutate(
    sample_global= sum(sample_strata)) %>% 
  left_join(ref_pop %>% 
              rename(pop_strata= "total_hh")) %>% 
  mutate(pop_global=sum(pop_strata) ,
  weights = (pop_strata/pop_global)/(sample_strata/sample_global)
  )


```



